Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM

### 2011-02-10 intro.tex
Dimensionless conversion... start

The Nernst-Planck equation for a mobile species ---
in our case for counter ions --- has the form
\begin{equation}
  \frac{\partial C}{\partial t}+\nabla\cdot(-D\nabla C-z\mu FC\nabla\phi)=0.
  \label{eq:nernst-planck}
\end{equation}
Here $C$ stands for the counter ion concentration, $D$ is diffusion, $\mu$ mobility,
$F$ Faraday constant, $\phi$ voltage, and $z$ the charge number. We have neglected 
the velocity of the species as in our case it can be assumed zero. 
The Poisson equation has the form
\begin{equation}
  -\nabla^2\phi=\frac{F\rho}{\varepsilon}
  \label{eq:poisson}
\end{equation}
where $\varepsilon$ is the absolute dielectric permittivity. The
charge density $\rho = C-C_{0}$ 
where $C_{0}$ is a constant anion concentration.

>>>>

The Nernst-Planck equation for a mobile species ---
in our case for counter ions --- has the form
\begin{equation}
  \frac{\partial C}{\partial t}+\nabla\cdot(-D\nabla C-\mu FC\nabla\phi)=0.
  \label{eq:nernst-planck}
\end{equation}
Here $C$ stands for the counter ion concentration, $D$ is diffusion, $\mu$ mobility,
$F$ Faraday constant, and $\phi$ voltage. We have neglected 
the velocity of the species as in our case it can be assumed zero. 
The Poisson equation has the form
\begin{equation}
  -\nabla^2\phi=\frac{F\rho}{\varepsilon}
  \label{eq:poisson}
\end{equation}
where $\varepsilon$ is the absolute dielectric permittivity. The
charge density $\rho = C-C_{0}$ 
where $C_{0}$ is a constant anion concentration.

### 2011-02-10 model.tex
Dimensionless conversion... start

As there is no flow through the domain's boundary, Eq.~\eqref{eq:nernst-planck}
is equipped with a Neumann boundary condition 
\begin{equation}
  -D \frac{\partial C}{\partial n} - z \mu F C \frac{\partial \phi} {\partial n} = 0.
  \label{eq:nernst-planck-boundary}
\end{equation}
Furthermore, we prescribe a positive constant voltage $V_{pos}$ 
on $\Omega_1$ and zero voltage on $\Omega_3$:
\begin{eqnarray}
  \phi_{\partial\Omega_1}&=&V_{pos},\\
  \phi_{\partial\Omega_3}&=&0.
  \label{eq:dirichlet}
\end{eqnarray}
On the rest of the boundary, $\phi$ has zero normal derivatives, and thus we prescribe 
a Neumann boundary condition
\begin{equation}
  \frac{\partial \phi_{\Omega_2}}{\partial n}=\frac{\partial \phi_{\Omega_4}}{\partial n}=0.
\end{equation}

@@

% IF YOU WANT TO CITE THIS WORK< YOU HAVE TO SAY WHY WE ARE DOING 
% THE DERIVATION AGAIN -- OTHERWISE IT IS CONTRAPRODUCTIVE. SO 
% I COMMENTED THIS OUT.
% I guess it can be left out as the work in SPIE is quite preliminary
% and not that complete
%An implicit version of the weak form of PNP was derived in~\cite{pugal2010spie}.
To make our results easily reproducible, in the following we present the derivation 
of weak forms of Eqs.~\eqref{eq:nernst-planck} and 
\eqref{eq:poisson}, as well as formulas for the Jacobian matrix and residual 
vector that are used in actual computations.
To simplify notation, we define
\begin{equation}
  K = z \mu F,\ \ \
  L = \frac{F}{\varepsilon}.
  \label{eq:KL}
\end{equation}
Eqs.~\eqref{eq:nernst-planck} and \eqref{eq:poisson} yield
\begin{eqnarray}
  \frac{\partial C}{\partial t}-D\nabla^2 C-K\nabla\cdot \left(C\nabla\phi\right)&=&0,\label{eq:nernst-planck-2}\\
  -\nabla^2\phi-L\left(C-C_{0}\right)&=&0.\label{eq:poisson-2}
\end{eqnarray}
Boundary condition Eq.~\eqref{eq:nernst-planck-boundary} has the form
\begin{equation}
  -D\frac{\partial C}{\partial n}-KC\frac{\partial\phi}{\partial n}=0.
  \label{eq:nernst-planck-boundary-2}
\end{equation}
As the second derivatives of both $C$ and $\phi$ are present in the 
equations, the appropriate function space for them is the Sobolev space 
$V=H^1\left(\Omega\right)$ where 
$$H^1\left(\Omega\right)=\left\{v\in L^2\left(\Omega\right);\ \nabla v \in \left[L^2\left(\Omega\right)\right]^2\right\}.
$$
In order to derive the weak form of the Nernst-Planck equation Eq.~\eqref{eq:nernst-planck-2},
we first multiply it with a test function $v^C \in V$ and integrate over the domain~$\Omega$,
\begin{equation}
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}
  -\int_{\Omega}D\nabla^2Cv^C d\mathbf{x}-\int_{\Omega}K\nabla C\cdot\nabla\phi v^C d\mathbf{x}
   - \int_{\Omega}KC\nabla^2\phi v^C d\mathbf{x}=0.
  \label{eq:nernst-planck-weak1}
\end{equation}
Applying the Green's first identity to the terms that contain second derivatives,
we obtain
$$
 \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C\cdot\nabla v^C d\mathbf{x}-
  K\int_{\Omega}\nabla C \cdot \nabla \phi v^C d\mathbf{x}
$$
\begin{equation}
  + K\int_{\Omega}\nabla\left(Cv^C\right)\cdot \nabla \phi d\mathbf{x}
  -D\int_{\partial\Omega}\frac{\partial C}{\partial n}v^C d\mathbf{S}-
  \int_{\partial\Omega}K\frac{\partial\phi}{\partial n}Cv^C d\mathbf{S}=0.
  \label{eq:nernst-planck-weak2}
\end{equation}
Expanding the nonlinear term and using the boundary condition 
\eqref{eq:nernst-planck-boundary-2}, we have
$$
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C \cdot \nabla v^C d\mathbf{x}-
  K\int_{\Omega}\nabla C \cdot \nabla \phi v^C d\mathbf{x}
$$
\begin{equation}
  + K\int_{\Omega}\nabla \phi \cdot \nabla C v^C d\mathbf{x}+
  K\int_{\Omega} C \left(\nabla\phi\cdot\nabla v^C\right) d\mathbf{x}=0.
  \label{eq:nernst-planck-weak3}
\end{equation}
After the second and third terms cancel out, we obtain the final weak form of 
the Nernst-Planck equation
\begin{equation}
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C \cdot \nabla v^C d\mathbf{x}+
  K\int_{\Omega} C \left(\nabla\phi\cdot\nabla v^C\right) d\mathbf{x}=0.
  \label{eq:nernst-planck-weak-final}
\end{equation}
Analogously we derive also the weak form of the Poisson equation (\ref{eq:poisson-2}),
\begin{equation}
  -\int_{\Omega}\nabla^2\phi v^\phi d\mathbf{x}-\int_{\Omega}LCv^\phi d\mathbf{x}+
  \int_{\Omega}LC_{0}v^\phi d\mathbf{x}=0.
  \label{eq:poisson-weak1}
\end{equation}
Performing integration by parts and taking into account the boundary 
conditions for $\phi$, we obtain
\begin{equation}
  \int_{\Omega}\nabla\phi\cdot\nabla v^\phi d\mathbf{x}-\int_{\Omega}LCv^\phi d\mathbf{x}+
  \int_{\Omega}LC_{0}v^\phi d\mathbf{x}=0.
  \label{eq:poisson-weak-final}
\end{equation}

>>>>

As there is no flow through the domain's boundary, Eq.~\eqref{eq:nernst-planck}
is equipped with a Neumann boundary condition 
\begin{equation}
  -D \frac{\partial C}{\partial n} - \mu F C \frac{\partial \phi} {\partial n} = 0.
  \label{eq:nernst-planck-boundary}
\end{equation}
Furthermore, we prescribe a positive constant voltage $V_{pos}$ 
on $\Omega_1$ and zero voltage on $\Omega_3$:
\begin{eqnarray}
  \phi_{\partial\Omega_1}&=&V_{pos},\\
  \phi_{\partial\Omega_3}&=&0.
  \label{eq:dirichlet}
\end{eqnarray}
On the rest of the boundary, $\phi$ has zero normal derivatives, and thus we prescribe 
a Neumann boundary condition
\begin{equation}
  \frac{\partial \phi_{\Omega_2}}{\partial n}=\frac{\partial \phi_{\Omega_4}}{\partial n}=0.
\end{equation}

@@

To make our results easily reproducible, in the following we present the derivation 
of weak forms of Eqs.~\eqref{eq:nernst-planck} and 
\eqref{eq:poisson}, as well as formulas for the Jacobian matrix and residual 
vector that are used in actual computations.
To simplify notation, we use dimensionless formulation of Eqs.~\eqref{eq:nernst-planck}
and~\eqref{eq:poisson}. The following new notations for the independent variables
$x$, $y$, $t$ and for the dependent variables $C$ and $\phi$ are used~\cite{bazant2004diffuse}:
\begin{equation}
  X=\frac xL,\ \ \
  Y=\frac yL,\ \ \
  \tau=\frac{tD}{\lambda_D L},\ \ \
  \varphi=\frac{\phi F}{RT},\ \ \
  c=\frac C C_0.\label{eq:dimensionless}
\end{equation}
Here $\lambda_D$ is the Debye screening length and it is expressed as follows~\cite{bazant2004diffuse}:
\begin{equation}
  \lambda_D=\sqrt{\frac{\varepsilon R T}{2 F^2 C_0}}.\label{eq:debye}
\end{equation}
After inserting variables~\eqref{eq:dimensionless} into Eq.~\eqref{eq:nernst-planck} the Nernst-Planck
equation and Poisson equation become:
\begin{eqnarray}
  \frac{DC_0}{\lambda_D L}\frac{\partial c}{\partial \tau} + \frac 1L \nabla_d\cdot
    \left(-\frac{DC_0}{L}\nabla_dc-c\frac{DC_0}{L}\nabla_d\varphi\right) &=& 0\label{eq:dim1}\\
  - \frac{\varepsilon R T}{L^2F^2C_0}\nabla_d\varphi&=&c-1\label{eq:dim2},
\end{eqnarray}
where
\begin{equation}
  \nabla_d=\left(\frac{\partial}{\partial X},\ \frac{\partial}{\partial Y}\right).
\end{equation}
After simplifying Eqs.~\eqref{eq:dim1} and~\eqref{eq:dim2} and denoting
\begin{equation}
  \epsilon=\frac{\lambda_D}{L},
\end{equation}
the dimensionless form of the PNP system of equations is
\begin{eqnarray}
  \frac{\partial c}{\partial \tau} + \epsilon \nabla_d\cdot
    \left(-\nabla_d c - c \nabla_d \varphi\right) &=&0,\label{eq:dimensionless-nernst-planck}\\
  -2\epsilon^2\nabla_d\varphi&=&c-1\label{eq:dimensionless-poisson}.
\end{eqnarray}
Boundary condition Eq.~\eqref{eq:nernst-planck-boundary} has the form
\begin{equation}
  -\frac{\partial c}{\partial n}-c\frac{\partial\varphi}{\partial n}=0.
  \label{eq:dimensionless-nernst-planck-boundary}
\end{equation}
As the second derivatives of both $c$ and $\varphi$ are present in the 
equations, the appropriate function space for them is the Sobolev space 
$V=H^1\left(\Omega\right)$ where 
$$H^1\left(\Omega\right)=\left\{v\in L^2\left(\Omega\right);\ \nabla v \in \left[L^2\left(\Omega\right)\right]^2\right\}.
$$
In order to derive the weak form of the Nernst-Planck equation Eq.~\eqref{eq:dimensionless-nernst-planck},
we first multiply it with a test function $v^c \in V$ and integrate over the domain~$\Omega$,
\begin{equation}
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}
  -\int_{\Omega}D\nabla^2Cv^C d\mathbf{x}-\int_{\Omega}K\nabla C\cdot\nabla\phi v^C d\mathbf{x}
   - \int_{\Omega}KC\nabla^2\phi v^C d\mathbf{x}=0.
  \label{eq:nernst-planck-weak1}
\end{equation}
Applying the Green's first identity to the terms that contain second derivatives,
we obtain
$$
 \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C\cdot\nabla v^C d\mathbf{x}-
  K\int_{\Omega}\nabla C \cdot \nabla \phi v^C d\mathbf{x}
$$
\begin{equation}
  + K\int_{\Omega}\nabla\left(Cv^C\right)\cdot \nabla \phi d\mathbf{x}
  -D\int_{\partial\Omega}\frac{\partial C}{\partial n}v^C d\mathbf{S}-
  \int_{\partial\Omega}K\frac{\partial\phi}{\partial n}Cv^C d\mathbf{S}=0.
  \label{eq:nernst-planck-weak2}
\end{equation}
Expanding the nonlinear term and using the boundary condition 
\eqref{eq:nernst-planck-boundary-2}, we have
$$
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C \cdot \nabla v^C d\mathbf{x}-
  K\int_{\Omega}\nabla C \cdot \nabla \phi v^C d\mathbf{x}
$$
\begin{equation}
  + K\int_{\Omega}\nabla \phi \cdot \nabla C v^C d\mathbf{x}+
  K\int_{\Omega} C \left(\nabla\phi\cdot\nabla v^C\right) d\mathbf{x}=0.
  \label{eq:nernst-planck-weak3}
\end{equation}
After the second and third terms cancel out, we obtain the final weak form of 
the Nernst-Planck equation
\begin{equation}
  \int_{\Omega}\frac{\partial C}{\partial t}v^C d\mathbf{x}+
  D\int_{\Omega}\nabla C \cdot \nabla v^C d\mathbf{x}+
  K\int_{\Omega} C \left(\nabla\phi\cdot\nabla v^C\right) d\mathbf{x}=0.
  \label{eq:nernst-planck-weak-final}
\end{equation}
Analogously we derive also the weak form of the Poisson equation (\ref{eq:poisson-2}),
\begin{equation}
  -\int_{\Omega}\nabla^2\phi v^\phi d\mathbf{x}-\int_{\Omega}LCv^\phi d\mathbf{x}+
  \int_{\Omega}LC_{0}v^\phi d\mathbf{x}=0.
  \label{eq:poisson-weak1}
\end{equation}
Performing integration by parts and taking into account the boundary 
conditions for $\phi$, we obtain
\begin{equation}
  \int_{\Omega}\nabla\phi\cdot\nabla v^\phi d\mathbf{x}-\int_{\Omega}LCv^\phi d\mathbf{x}+
  \int_{\Omega}LC_{0}v^\phi d\mathbf{x}=0.
  \label{eq:poisson-weak-final}
\end{equation}
